---
title: "ClientReport_MixedModel"
author: "ESRA TURKHAN s3411893"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Please adjust the directory according to your file location before running the code.
```{r echo=TRUE}
setwd("/Users/esrakasap/Desktop/adsız klasör/Statistical Consulting/Client Meetings")

#install.packages("readxl")
library(readxl)


df = read_excel("Tree_Inoculation_Clean.xlsx" , sheet = 2)

```

# To see the data structure initially, let's take a look at the first 10 rows of our dataset.
```{r echo=TRUE}
head(df)
```
#Obtain a statistical summary for each column in our dataset.
```{r echo=TRUE}
df$leaves <- as.numeric(df$leaves)
summary(df)

```
# Examine the distribution of leaf counts in each week across all datasets.
```{r echo=TRUE}


#install.packages("dplyr")
library(dplyr)

weekly_data <- df %>% group_by(measurement) %>% select(leaves)

library(ggplot2)


ggplot(weekly_data, aes(x = leaves)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  facet_wrap(~ measurement, scales = "free") +
  labs(title = "Leaves Distribution by Week", x = "Number of Leaves", y = "Frequency")

```
#Examine the distribution of leaf counts in each week across all datasets.
```{r echo=TRUE}
ggplot(df, aes(x = as.factor(measurement), y = leaves)) +
  geom_boxplot(fill = "blue", color = "black") +
  labs(title = "Boxplot of Leaves by Week", x = "Week", y = "Number of Leaves") +
  theme_minimal()
```
#Investigate how far the leaf count distributions deviate from a normal distribution.
```{r echo=TRUE}
ggplot(df, aes(sample = leaves)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~measurement, scales = "free") +
  labs(title = "QQ Plot of Leaves by Week")
```

#Check the missing values
```{r echo=TRUE}
missing_values = is.na(df)
missing_count = colSums(missing_values)
missing_count = as.data.frame(missing_count)
print(missing_count ,n = 20)

```
# more detailed investigation on NA s in the Leaves column.
```{r echo=TRUE}

missing_count_in_Leaves <- sum(is.na(df$leaves))
cat("Number NA in Leaves column:", missing_count_in_Leaves, "\n")

#The number of NA values in the Leaves column when the Mortality column is 1.
missing_count_in_Leaves_Mortality1 <- sum(is.na(df$leaves[df$mortality == 1]))
cat("Number of NA when Mortality=1:", missing_count_in_Leaves_Mortality1, "\n")

#The number of NA values in the Leaves column when the Mortality column is 0.
missing_count_in_Leaves_Mortality0 <- sum(is.na(df$leaves[df$mortality == 0]))
cat("Number of NA when Mortality=0 :", missing_count_in_Leaves_Mortality0, "\n")

which(df$mortality == 0 & is.na(df$leaves))
filtered_data <- df[df$mortality == 0 & is.na(df$leaves), ]
print(filtered_data)

```

#Remove the columns taht mortality=0 and Leaves=NA 
```{r echo=TRUE}

df <- filter(df, !(mortality == 0 & is.na(leaves)))

```


#Mean of leaves counts per week for each isolate
```{r echo=TRUE}


Leaves_mean <- aggregate(leaves ~ treatment + isolate + measurement + SET, data = df, mean)

print(Leaves_mean)

````

# Weekly change on the leaves count in set3
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==3), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(5, 8), color = "red", size = 0.5) +
  
  
  labs(title = "Spaghetti Plot -SET3-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```



# Weekly change on the leaves count in set4
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==4), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET4-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```
# Weekly change on the leaves count in set5
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==5), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET5-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```
# Weekly change on the leaves count in set6
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==6), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET6-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```
# Weekly change on the leaves count in set7
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==7), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET7-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```
# Weekly change on the leaves count in set8
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==8), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET8-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```


# Weekly change on the leaves count in set9
```{r echo=TRUE}
# ggplot2 paketini yükle
library(ggplot2)



# ggplot ile spaghetti plot'u oluştur
ggplot(Leaves_mean[which(Leaves_mean$SET==9), ], aes(x = measurement, y = leaves, color = isolate, linetype = treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(~isolate, scales = "free_y") +
  
  geom_vline(xintercept = c(4, 10), color = "red", size = 0.5) +
  
  labs(title = "Spaghetti Plot -SET9-",
       x = "Measurement Time",
       y = "Leaf Count",
       color = "Isolate",
       linetype = "Treatment")


```
# Fill the NA with 0 in the Leaves column.
```{r echo=TRUE}

df$leaves[is.na(df$leaves)] <- 0

```

#Filter the drought treatment weeks per each set
```{r echo=TRUE}
#install.packages("dplyr")
library(dplyr)


df_new <- df %>%
  filter((SET == 3 & measurement %in% c(5, 6, 7, 8)) |
         (SET != 3 & measurement %in% c(4, 5, 6, 7, 8, 9, 10)))


```

#Inversigate the distribution of leaves count during the drought treatment
```{r echo=TRUE}
#hist(df$leaves, main = "Distribution of Leaves", xlab = "Number of Leaves")

#install.packages("dplyr")
library(dplyr)

weekly_data_new <- df_new %>% group_by(measurement) %>% select(leaves)

library(ggplot2)


ggplot(weekly_data_new, aes(x = leaves)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  facet_wrap(~ measurement, scales = "free") +
  labs(title = "Leaves Distribution by Week", x = "Number of Leaves", y = "Frequency")

```
# Apply mixed model (isolate and treatment are fixed effects. SET is a rabdom effect)
```{r echo=TRUE}
#install.packages("lme4")
#install.packages("lmerTest")
library("lmerTest")
library(lme4)


lmer_model_leaves <- lmer(leaves ~ isolate * treatment + (1 | SET) , data = df_new)
summary(lmer_model_leaves)

fixed_effect_table = summary(lmer_model_leaves)$coefficients

```





# Order the isolte : 1) With respect to amount of their contribution on drougt resistance : 2) With respect to their significance
```{r echo=TRUE}
fixed_effect_table = as.data.frame(fixed_effect_table)
isolate_table_leaves = head(fixed_effect_table, 73)
top_rows_estimate_leaves <- isolate_table_leaves[order(isolate_table_leaves$Estimate, decreasing = TRUE), ][1:14, ]
top_rows_pValue_leaves <- isolate_table_leaves[order(isolate_table_leaves$`Pr(>|t|)`, decreasing = FALSE), ][1:100, ]


top_rows_estimate_leaves = as.data.frame(top_rows_estimate_leaves)
print(top_rows_estimate_leaves,n=14)

print(top_rows_pValue_leaves)


```

#########  EXTRA ANALYSIS TO INVESTIGATE OTHER FEATURES LIKE HEIGHT - ROOT FEATURES - SPAD ###########


#HEIGHT ANALYSIS
```{r echo=TRUE}
#install.packages("lme4")
#install.packages("lmerTest")
library("lmerTest")
library(lme4)

df_new$height = as.numeric(df_new$height)
lmer_model_height <- lmer(height ~ isolate * treatment + (1 | SET) , data = df_new)
summary(lmer_model_height)

fixed_effect_table_height = summary(lmer_model_height)$coefficients

```

```{r echo=TRUE}
fixed_effect_table_height = as.data.frame(fixed_effect_table_height)
isolate_table_height = head(fixed_effect_table_height, 73)
top_rows_estimate_height <- isolate_table_height[order(isolate_table_height$Estimate, decreasing = TRUE), ]
top_rows_pValue_height <- isolate_table_height[order(isolate_table_height$`Pr(>|t|)`, decreasing = FALSE), ]


top_rows_estimate_height = as.data.frame(top_rows_estimate_height)
print(top_rows_estimate_height,n=14)

print(top_rows_pValue_height)


```







